---
title: "Data Wrangling Notes"
author: "Claire Brandes"
date: "2024-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```
```{r}
install.packages("tidyverse")
library(tidyverse)
```


## Loading in practice data
```{r}
d <- read_csv("data-wrangling.csv", col_names = TRUE)
```

#### Note: read.csv function in base R, read_csv function in tidyverse with more features

## Building scientific name from two columns in data frame

```{r}
sciname <- paste(d$Genus, d$Species, sep = " ")
```

#### Another method of doing the same thing

```{r}
sciname <- paste0(d$Genus, " ", d$Species)
```

## Make everything in sciname uppercase

```{r}
sciname <- toupper(sciname)
```

## Pull out all of the unique genera in data set

```{r}
unique(d$Genus)
```

#### Another method of doing the same thing

```{r}
unique(d[["Genus"]])
```

# Ways to look at data frame

## Give structure of dataframe

```{r}
glimpse(d)
```

## Same information, but not as good as tidyverse version

```{r}
str(d)
```

## First 6 rows of data frame

```{r}
head(d)
```

## Last 6 rows of data frame

```{r}
tail(d)
```

## First 10 rows of dataframe

```{r}
head(d, 10)
```

# Manipulating data set

```{r}
gs <- d$MeanGroupSize
```

## Add 1 to each number in vector

```{r}
gs + 1
```

```{r warning=TRUE}
gs + c(1,2)
```
# Extract female mean body mass and brain size, and make a bivariate plot of the relationship between those variables on untransformed and log scale 

## Name variables 
```{r}
bm <- d$Body_mass_female_mean
bs <- d$Brain_Size_Species_Mean
```
## Create untransformed plot
```{r}
plot(bm,bs)
```
## Create transformed plot
```{r}
plot(log(bm),log(bs))
```


